---
title: "Intro to Character Strings"
author: "Jamie Reilly"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: paged
    css: jamie_mod4.css
---

```{r setup, include=FALSE, cache=T}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', echo=TRUE, tidy=TRUE, message=F, warning=F, cache=T)
```

Libraries 
```{r echo=T, message=F}
library(tidyverse)
jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())
```

# Character strings
```{r}
#close characters with quotes or else they get executed as code
a <- "hello world!"
print(a)
```

# Creating character vectors
```{r}
planets <- c("m", "v", "e", "m", "j", "s", "u", "n", "p")  #separate elements
planets
typeof(planets)  #test type of vector 4 ways
is.character(planets) #boolean
class(planets)
str(planets)
```

# Strings as factors
By default, character strings inside a data frame will be converted to factors. You need to tell R explicitly not to do this using stringsAsFactors = FALSE
```{r}
dfA <- data.frame(numbers=1:5, letters=letters[1:5])
dfB <- data.frame(numbers=1:5, letters=letters[1:5], stringsAsFactors = FALSE)
```

# Paste
The paste command is super useful for characters. It takes this form
paste(..., sep = " ", collapse = NULL) -- where the default seperator is a space.  The arguments get executed as character strings. Here's an example combining a statement with printing the system date. 

```{r}
saythedate <-  paste("Hi! today is", (Sys.Date()))
saythedate
```

# Paste0
Is equivalent to collapse = T
```{r}
a <- paste0("get", "me", "the", "hell", "out", "of", "here")
b <- paste0("get me the hell out of here") #this doesn't work!
print(a)
print(b)

```

# Letters
## seperators

```{r}
cat(2:12, sep = "+")  #means concatenate and print
cat(2:12, sep = "/") 
cat(2:12, sep = "a") 
```

## Case manipulation
```{r}
dat <- c("ThE Day wAS LONG and TeRriBle.")  #convert to lowercase
dat.l <- tolower(dat)
print(dat.l)
```

## string splitting
strsplit(string, delimiter)
```{r}
strsplit(dat.l, " ") %>% print()
```

# String pattern matching
## grep
base R returns indices of matching arguments

```{r}
try <- c("a", "b", "a", "c", "z", "b", "a")
grep(try)
```


length(colours())


