---
title: "Working with characters: Part I - In Class"
author: "Ann Marie Finley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    df_print: paged
    css: jamie_mod5.css
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', echo=TRUE, tidy=TRUE,  cache=T, message=F, warning=F)
```

```{r echo=F}
library(tidyverse)
```

# Characters & Strings
```{r}
a <- "hello world!"; a #enclose characters with quotes or else they get executed as code
```

Generate a count of the characters in a string or vector
```{r}
nchar(a) #counts the number of characters including spaces
```

## Strings 
Strings are sequences of more than one characters
```{r, eval=T}
planets <- c("m", "v", "e", "m", "j", "s", "u", "n", "p"); planets  #separate elements
is.character(planets) #boolean evaluates to True (1) or False (0) - you can count the number of a condition
```

## Paste
Converts data to character and concatenates into vectors separating by the sep argument.  <br />
It takes this form paste(..., sep = " ", collapse = NULL) -- where the default separator is a space.  <br />
Here's an example combining a statement with printing the system date. If you specify collapse=T, you will remove all delimiters (spaces here)
```{r, eval=T}
paste("Hi! today is", (Sys.Date()))  #collapse=NULL
```

## Separators & delimiters
'cat' concatenates and prints a character vector
```{r, eval=T}
cat(2:12, sep = "+")  #the '+" is a delimiter here
cat(2:12, sep = "/") 
cat(2:12, sep = "a") 
```

## Case Manipulation
```{r, eval=T}
dat <- c("ThE Day wAS LONG and TeRriBle."); dat  
dat.low <- dat %>% tolower() #convert to lowercase
dat.up <- dat %>% toupper() #convert to uppercase
print(dat.low)
print(dat.up)
```

## String Splitting
Sometimes you need to split the elements of a string (e.g., tokens) into separate units. <br />
The strsplit(string, delimiter) command is useful for this. This takes the form: <br />
newdat <- strsplit(olddat, "separator [usually a space] ") - turns things into a LIST

```{r, eval=T}
length(dat.low)  #print 
splitup <- strsplit(dat.low, " ") %>% print()
length(splitup)
new <- c("hi i'm not feeling well"); new #we need to split/tokenize vector by word
snew <- strsplit(new, " "); snew #the inside of the quotes marks how the different tokens are divided
```
<br />

# Detect, Locate, Extract, Replace String Patterns
## Detect
Grep - Is a pattern happening?

## Locate
regexpr(pattern, string) - see other markdown <br />
find starting position and length of first match

## Extract 
regmatches(string, regexpr(pattern, string))
extract first match [1] "tam" "tim"

## Replace
sub(pattern, replacement, string)
replace first match - stringr pattern does a similar thing
<br /> gsub = a global search and replace
<br />
<br />

# Grep & grepl
We are going to work a bit with a built in character vector called colors().
```{r}
head(colors(), n=25) #colors is a built in character vector of color palettes in R
length(colors())
typeof(colors())
str(colors())
tail(colors(), n=3)
(colors()[329]) #brakets say "give me row 329 - if you have nothing after, it'll return ALL associated cols. brackets essentially say [row, col]
(colors()[505:508]) #give me rows 505 to 508; R likes : for sequences - this is subsetting; similar to slice


```

Base R default returns indices of matching arguments -- [1] 1st observation column <br />
The grep function format involves pattern match where pattern is a character string <br />
grep(pattern, x, ignore.case = FALSE, perl = FALSE, value = FALSE,
     fixed = FALSE, useBytes = FALSE, invert = FALSE)

```{r}
try <- c("a", "b", "a", "c", "z", "b", "a")
grep("a", try, value=F) #will give indices - each row index of each occurrence, numerically
grep("a", try, value=T) #will give values for words with an 'a' in them - what words have an "a" in them here?
grep("a", try, value=T, invert=T) #inverse -- returns words WITHOUT an 'a' in them
```

Let's try this with a much longer character vector --  colors()
```{r}
grep("green", colors(), value=F) #will give indices with the string green in them
grep("green", colors(), value=T) #will give any word with "green" in it


blue <- grep("blue", colors(), value=F) #make a set of blue colors
length(blue) #how many colors have blue in them?
```

Let's inspect a few indices -- maybe 85 to 88
```{r}
colors()[85:88]
```

How many shades of green are in the R color palette?
```{r}
length(grep("green", colors(), value=F))
```

Now let's look for indices that have the words "green" or "white" in them; | means "or"
```{r}
grep("green|white", colors(), value=T) #help jamie

borange <- grep("orange", colors(), value=T, invert=T) #what colors don't have orange
length(borange)

graqua <- grep("green|aqua", colors(), value=T)  #what colors have green or aqua
length(graqua)

```
<br />

Let's return values instead of indices
```{r}
grep("green", colors(), value=F) #will give values with the string green in them
```

# Grepl
returns a boolean true/false at each index
```{r}
grepl("green", colors()) #can sum truths to get count of green words
```

# Setdiff
contrasts two vectors, subtract out elements/values in one from another -- think of subtracting brain 1 activation from brain 2 activation 
```{r}
me <- c("philadelphia", "new york", "boston")
you <- c("philadelphia", "new york", "boston", "new orleans")
setdiff(me, you)  #me versus you what is different? --> there is nothing different, everything in me is in you
setdiff(you, me)  #you versus me what is different? --> new orleans is different

set.seed(123)
a <- as.data.frame(sample(letters[1:26], 200, replace=T))
head(a)

set.seed(456)
b <- as.data.frame(sample(letters[1:26], 200, replace=T))
head(b)

setdiff(a, b)
```

# Sub & gsub
substitutes one string for another gsub= global substutition - gsub(take this, relplacethat, invector)
```{r, eval=F}
you<- c("you make me sick!", "you really do.")   #sub(pattern = "x", replacement = "m", vector name)
print(you)
me <- sub("you", "I", you)  #only does first instance
print(me)
globe <- gsub("you", "I", you) #global - substututes all instances
print(globe)

```

# Try Substituting Colors
```{r}
colors1 <- colors()
length(grep("blue", colors1, value=T))  #how many colors have blue

colors2 <- gsub("blue", "cat", colors1) #substitute cat for blue
length(grep("blue", colors2, value=T))  
length(grep("cat", colors2, value=T))  
```


