---
title: "TU_TextMine_Midterm"
author: "Ann Marie Finley"
date: "9/25/2020"
output:
  html_document:
    toc: true
    toc_depth: 1
    df_print: paged
    css: jamie_mod6.css
---
```{r setup, include=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', eval=T, echo=T, tidy=TRUE,  cache=T, message=F, warning=F)  
```

```{r echo=F}
library(tidyverse)
library(stringi)
library(wordcloud)
library(corpus)
library(RColorBrewer)
library(tm)  
library(qdapRegex)
library(mgsub)
```

# Introduction: We Should All Be Feminists
## By Chimimanda Ngozi Adiche
XX why this project, significance to me, interesting - what do I think? Feminist is pos or neg? Lots of cleaning bc of punctuation, time marks  <br />

# Reading in the Text
```{r}
fem <- readLines("feminists_adiche.txt") #Read in the plain text file
str(fem) #Examine structure of data
print(fem) #Inspect text file
```

There is a lot to clean here:
1. Spacing
2. Punctuation: watch out for intra-word dashes and quotes
3. Odd character strings
4. Annotations 
5. Digits 
6. Case <br />

# Cleaning
## Shaping a Concatenated Character Vector - Spacing
```{r}
fem1 <- paste(readLines("feminists_adiche.txt"), collapse = " ") #Reformat by concatenating into a single row by gluing all spacing together.
str(fem1) #Examine structure of data
length(fem1) #I'm anal about checking the number of rows
print(fem1) #Spacing has been fixed
#fem1 %>% stri_wrap() %>% print() #Reprint so that it does not break across words, this is aesthetics only - it worked a couple times and now does not?!****************
```

So far so good! Let's start to think about some of the tricker aspects of cleaning.  <br />

## Removing Punctuation
```{r}
mgsub <- function(pattern, replacement, x, ...) {
  if (length(pattern)!=length(replacement)) {
    stop("pattern and replacement do not have the same length.")
  }
  result <- x
  for (i in 1:length(pattern)) {
    result <- gsub(pattern[i], replacement[i], result, ...)
  }
  result
}

fem2 <- gsub('[[:punct:] ]+',' ', fem1) #Strip out punctuation - But this smooshed the entire paragraph together. This will work nicely once character is vectorized and unlisted etc - otherwise it's word vomit.

fem3 <- removePunctuation(fem1, preserve_intra_word_contractions = F); fem3 #try another way of stripping the punctuation. This seemed to work okay. I still have that odd character string. 

```

Onward! Now I need to edit out the digits and the annotations (Laughter).

## Removing Odd Character String
```{r}
fem4 <- gsub("\\xca", " ", fem3); fem4 #Strip out the odd pattern I noted in the visually inspected data. Replace it with a space. #Well, that is not what I wanted to happen. Gotta look at my gsub again. 

#Maybe multiple global substitutions will fix our problem. I had a lot of issues getting this code to run. I found this solution in the Kwartler text but then mgsub was experiencing a variety of issues.  

patterns <- c("\xca")
replacements <- c(" ")
fem5 <- mgsub(patterns,replacements, fem3); fem5 #this worked with the little intro code thing at the top to define mgsub but still not sure why it isn't loading in the package I installed. Even when I call in the respective libraries up top and re-run it breaks. But, it worked with this trick. 
```


## Removing Digits and Annotations
```{r}
fem6 <- str_split(fem5, " ") %>% unlist()
head(fem6)
str(fem6)
grep("[[:digit:]]", fem6, value=T) #Return all strings with digits in them
fem7 <- gsub("[[:digit:]]", "", fem6); fem7 #Replace numbers with nothing

grep("[Ll]aughter|[Aa]pplause", fem7, value=T) #Find annotations
gsub("[Ll]aughter|[Aa]pplause", "", fem7) #Replace annotation with nothing

fem8 <- stri_omit_empty(fem7, na_empty = T); fem8 #Remove empty tokens
str(fem8)

```

## Change to Lowercase
```{r}
tolower(fem8)
```


