---
title: "TU Datamining Week 2: Dataframes and Types"
author: "Ann Marie Finley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: paged
    css: jamie_mod4.css
---

```{r setup, include=FALSE, cache=T}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', echo=TRUE, tidy=TRUE, message=F, warning=F, cache=T)
```

# Libraries 
```{r echo=T, message=F}
library(tidyverse)
library(janitor)
library(tidyverse)
```

# Vectors
Uses the c() command. Contain all the same type of variable (integer, character, etc). Combines elements into a vector. 
```{r}
emma <- c("a", "b", "c") #strings are combinations of more than one character
print(emma)
typeof(emma) #tells me what type of data I am working with
```

Making a new vector
```{r}
amf <- seq(20, 25) #produce sequence 20-25 
print(amf) #figure out getting a series of numbers. 
#******Why does "by=1" change to "double" type instead of "vector" type********
typeof(amf) #tells me what type of data I am working with
```


# Dataset 1
We'll read in and inspect a very simple dataframe (dat_simple) here.  
```{r}
easy <- read.csv("dat_simple.csv", header=T)  #read in data, header is first row
#remember no spaces, csv or text is best, or CamelText
```

Look at its structure and check for missing observations
```{r}
str(easy) #str stands for structure. Here, I am checking the structure of easy
```

Print the whole dataframe
```{r}
print(easy) #look at whole dataframe, can also just type "easy" no quotes
```

Examine just the first seven rows
```{r}
head(easy, 7)
```

Examine just the last three rows
```{r}
tail(easy, 3)
```

What is the smallest and largest shoe size?
```{r}
max(easy$Shoe) #easy is the dresser, the drawer is Shoe, find max value
min(easy$Shoe) #easy is the dresser, the drawer is Shoe, find min value
range(easy$Shoe) #gives range of min to max values of Shoe within easy
```

How many IQ observations in my dataset?
```{r}
length(easy$IQ) #How long is this vector IQ in the dataframe easy?
#AKA how many IQ observations do I have? How does this compare to my total number of observations?
```


What is the average IQ?
```{r}
mean(easy$IQ) #mean value of IQ
summary(easy$IQ) #summary (descriptive) statistics on IQ
```

Scatterplot IQ by Shoe Size
```{r}
plot(easy$IQ, easy$Shoe, col="green") #simple scatterplot of IQ by Shoe Size
```

# Using the %>% function
This very helpful function "chains" commands so that you do not have to. Is called "piping" from dplyr package. Take a dataset - pipe it into the next command. Rather than "dresser and drawer" say "just look in this dresser," to R. 

newdataframe <- olddataframe %>% mutate(newvariablename = operation)

# ifelse function
ifelse(criteria, "something", "somethingelse")
If criteria, then something, if not then something else.
```{r}
str(easy) #inspect easy
easy2 <- easy %>% mutate(newvar = IQ-Shoe) #we made a new dataframe called easy2
easy2 #print DF

#now we want to group by big or little feet via a new variable
easy3 <- easy2 %>% mutate(BigFoot = ifelse(Shoe < 7,"small", "big") )
# Here, I mutate easy2 to easy3. I make a new var BigFoot, which queries the Shoe var in easy2 and categorizes them into small if less than 7 and big if anything else. BigFoot is a factor variable: i.e., do Bigger feet = bigger IQ?
easy3

easy4 <- easy3 %>% mutate(SmartorNo = ifelse(IQ <= 110, "No" , "Smart"))
easy4

easy5 <- easy4 %>% mutate(Weird = ifelse(Shoe*IQ <= 600 , "cherry","white"))  
easy5

#Check order of operations with parentheses around Shoce*IQ <=600
easy6 <- easy4 %>% mutate(Weird = ifelse((Shoe*IQ <= 600), "cherry","white"))  
easy6 #output is the same as easy5, but see "Square a Number"

```

square a number
```{r}
-3^2

```

# Subsetting a Variable
dplyr filter and select commands 
-- select out specific columns
```{r}
smallset <- easy4 %>% select(1:3) #grab columns 1-3 from  my original dataframe and write to a new object called # # smallset. Can also select(1, 5, 6)

#tinyset <- smallset %>% filter(1-4) *******#what is happening??? Jamie will see***************
```

# Transforming a variable
Create a new variable from using some operation on an old variable. 
```{r}

```

Is there any relationship between IQ and Shoe Size in this dataset?
```{r}

``` 


# Scaling up in complexity
Now we will look at some much more complex data (dat_complex) using many of the same scripting commands you used with the simple data.  This dataset represents lexical characteristics for almost 80,000 English words. 

# Read in new dataframe
```{r}
hard <- read.csv("dat_complex.csv" , header=T) %>% clean_names() #read in dataset and clean using janitor library
str(hard) #look at overall structure of hard dataframe

```

# Average frequency of a word
what is the average frequency per million words of an English word?
```{r}
hard %>% mean(brys_freqsubtlex, na.rm=TRUE) #********Why are you not working? JR will find out************8
mean(hard$brys_freqsubtlex,  na.rm=TRUE) #find average frequency, remove missing (NA) operations)
```

#Low and High Imageability
what is the lowest imageability and highest imageability word in English?
```{r}
#hard %>% range(gls_img, na.rm=TRUE)  # not working, stop here
```

what is the longest word in English?
```{r}

```

how many observations are in this dataset?
```{r}

```

data transformations and grouping variables
```{r}
#dat2 <- dat %>% mutate(newgroup, ifelse(nlet >=7, long, short))
```





# Extra credit challenge!
How many English words have more than 6 letters?
```{r}
# ifelse command
```

how to create a grouping variable -- are long words acquired later than short words?
```{r}
#lets say long words vs short words
```



