---
title: "Text Mining HW3: Frankenstein"
author: "Ann Marie Finley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 1
    df_print: paged
    css: jamie_mod6.css
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', eval=T, echo=T, tidy=TRUE,  cache=T, message=F, warning=F)  
```

```{r echo=F}
library(tidyverse)
```
<br />

# 1. Read in the data
Here's the first paragraph of Frankenstein by Mary Shelley  <br />

You will rejoice to hear that no disaster has accompanied the commencement of an enterprise which you have regarded with such evil forebodings.  I arrived here yesterday, and my first task is to assure my dear sister of my welfare and increasing confidence in the success of my undertaking.  <br >

write this paragraph to an object called blockhead. then print blockhead

```{r}
#write the para to blockhead
blockhead <- "You will rejoice to hear that no disaster has accompanied the commencement of an enterprise which you have regarded with such evil forebodings.  I arrived here yesterday, and my first task is to assure my dear sister of my welfare and increasing confidence in the success of my undertaking."; blockhead
```

what is the structure of blockhead?

```{r}
str(blockhead) #get structure
```

how many rows is blockhead?
```{r}
length(blockhead)
```

split (tokenize) blockhead into a vector of individual words. Call that bolthead
(hint use: str_split with " " as the separator)
```{r}
bolthead <- strsplit(blockhead, " "); bolthead #tokenize blockhead by word, delimiter is a space
str(bolthead) #check structure, it is a list of individual words within a single cell
bolthead1 <- unlist(bolthead) #pull all words into unique rows
str(bolthead1) #check structure

```

Convert bolthead to all upper case characters and print the first ten rows
```{r}
bolthead.up <- bolthead1 %>% toupper() #convert to uppercase
head(bolthead.up, n= 10)

```


How many characters are in blockhead?
```{r}
nchar(blockhead)

#I also look at bolthead1 here
nchar(bolthead1) #if you ask for nchar after unlisting, it gives you nchar per word/token e.g., by row

#what if we remove punctuation? Does R count punctuation as a characters?
blockhead1 <- gsub('[[:punct:] ]+',' ', blockhead) #here we removed punctuation and replaced with spaces
nchar(blockhead1) #R is still counting the spaces as characters

blockhead2 <- gsub('[[:punct:] ]+','', blockhead) #here we removed punctuation and replaced with [nothing]
nchar(blockhead2) #this is the number of letters (characters) 

```

How many words are in blockhead?
```{r}
lengths(gregexpr("[A-z]\\W+", blockhead)) + 1L #shoutout to google for this code. Another way to get this count would be to use the tokenized and unlisted character vector from bolthead1 as below

length(bolthead1)
```

Where did I go wrong in life?
```{r}
answer <- "In oH So mANy waYs." #write in answer
answer1 <- strsplit(answer, " ") #tokenize
answer2 <- unlist(answer1) #unlist
length(answer2) #how many words/rows
answer3 <- answer2 %>% tolower() #make it very ee cummings pt. 1
answer4 <- gsub('[[:punct:] ]+','', answer3) #make it very ee cummings pt. 2 - sub punctuation for '[nothing]'
tail(answer4, n = 3) #look at last three rows
```


