---
title: "Regular Expressions in R"
author: "Jamie Reilly. Ph.D."
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 1
    css: jr_v1.css
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', eval=T, echo=T, tidy=TRUE,  cache=T, message=F, warning=F)  
```

```{r echo=F}
library(tidyverse)
library(stringi)
```

# Regular Expressions
Regular expressions are also known as regex or regexp, and they are magical. Regular expressions involve a syntax for string matching of the sort used in find-and-replace algorithms. Here are some of the metacharacters we will be using in various text mining applications:  <br />
. | ( ) [ { ^ $ * + ?  - <br />

# Nomenclature of Metacharacters
["stuff_to_match"] brackets a character class-- a set of characters you hope to match with regular expressions. Make sure to honor the brackets and the quotes. <br />

### Description of metacharacters  <br />
'-' denotes a range or sequence such as ["a-d"] <br />
'.' denotes any character such as ["b.d"] (returns spaces, letters, etc.) <br />
'|' denotes 'or' (e.g., ["a|B"] return obs that have an a or a B in them <br />
'^' denotes 'not' (e.g.,["^9"] return obs that don't match 9 <br />
'\\' escapes a metacharacter <br />
'*' previous character repeated 0 or more times <br />
'+' previous character repeated 1 or more times 
<br />
<br />

### Help & practice
Regexp cheatsheet https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf <br />
Online regular expression checker https://spannbaueradam.shinyapps.io/r_regex_tester/ <br />
<br />

## More
Some Character Classes (copied from the RStudio cheatsheet):  <br />
1. [[:digit:]] or \\\\d digits; [0-9] <br />
2. \\\\D - non-digits; [^0-9] <br />
3. [[:lower:]] -- Lower-case letters; [a-z] <br />
4. [[:upper:]] -- Upper-case letters; [A-Z] <br />
5. [[:alpha:]] -- Alphabetic characters; [A-Za-z] <br />
6. [[:alnum:]] -- Alphanumeric characters [A-Za-z0-9] <br />
7. \\\\w -- Word characters; [A-Za-z0-9_] <br />
8. \\\\W -- Non-word characters <br />
9. [[:punct:]] -- Punctuation characters; !"#$%&’()*+,-./:;<=>?@[]^_`{|}~ <br />

Let's work with regular expressions on a familiar pangram with some extra junk in it. Read this phrase into an object called 'pan' so that each element separated by a space gets its own row in a character vector. Hint -- you need to unlist and/or split the string if you read it in raw.<br />

**The qU\*iccK brOwn Fox &^ Jump;s ove_r the Laz7y d'og .**

```{r}
#pan <-
#class(pan)
#print(pan)
```
Our goal is to eventually clean this up and eventually tokenize it. There are misspellings, errant punctuation, extra white space and a mix of upper and lowercase letters. Let's first check some of the observations using metacharacters and regular expressions. <br />
<br />

# One-by-one  
## . 
Let's grep our way to retrieving observations with some particular patterns where the form is grep(pattern, string, value=T). If you want to search or extract a character like "." that is also reserved as a metacharacter, you must escape the character with //  <br />
1. return all observations with an "i" followed by any character <br />
2. return all observations with a "T" followed by any character <br />
3. return all observations with a J followed by any character then an m <br />
4. return all observations where "^" is preceeded by any character
```{r}
#
#
#
#
```

## | means or
1. return all observations with an "i" or a "z" in them <br />
2. return all observations with an "o", "&", or "j" <br />
3. return all observations with a digit or an "F" <br />
4. return all observations of the pattern t _ e or  o _ e <br />
5. return all observations with a "z" and a "y" in them
```{r}
#
#
#
#
#
```

## [...] specifies permitted characters and [^...] not permitted characters
1. Return all observations with "the" or "The" 
2. Return all observations that do are not "The" or "the" 
```{r}
#
#
```

## [a-z] specifies a sequence of characters
1. Return all observations with letters a,b,c, or d in them
2. Return all observations with the numbers 7-9 in them
2. Return all observations that do not have the letters m-z in them
```{r}
#
#
```

# Character Classes
1. Return all entries with digits in them
2. Return all entries with punctuation in them
3. gsub out punctuation for an empty space -- form is gsub(pattern, relacement, vec) - write to pan2
```{r}
#
#
```

# Anchors
1. ^ Start of the string
2. $ End of the string
3. \\\\b Empty string at either edge of a word
4. \\\\B NOT the edge of a word
5. \\\\< Beginning of a word
6. \\\\> End of a word  <br />

Let's give some of these a shot
1. Return words in pan2 that start with a T or a t
2. Return words in pan2 that don't start with a T or a t
```{r}
#
#
```
<br />

# Quantifiers
1. \\* matches at most 0 times  <br />
2. \+ Matches at least 1 time  <br />
3. \? Matches at most 1 time; optional string <br />
4. \{n} Matches exactly n times  <br />
5. \{n,} Matches at least n times  <br />
6. \{,n} Matches at most n times <br />
7. \{n,m} Matches between n and m times <br />

let's try some quantifiers <br />
1. Return words that with at least two c's in them
```{r, eval=T}
#
```

# More advanced regexp practice
Let's see how some these expressions work in a mixed string of letters, numbers, and punctuation.
```{r, eval=T, echo=T}
set.seed(123) #stri_rand_strings(n, length, pattern = "[A-Za-z0-9]")
weird <- stri_rand_strings(25, sample(3:7), pattern = "[A-Za-z0-9;#$*@!]")  
print(weird)
str(weird)
```

# Exercises
These exercises index the vector "weird" you just created in the last step.
1. grep strings that have a vowel in them
2. grep strings that don't have a vowel 
3. grep strings that have an "a" or "p" anywhere in them.
4. grep strings that have are six characters long
5. grep strings that begin with a consonant
6. grep strings that begin with a digit
7. grep strings that end with a vowel
8. grep strings that have one or more vowels. 
9. grep strings that start with punctuation.
10. gsub punctuation for nothing.
11. convert all letters to lowercase
12. eliminate all numbers
13. strip whitespace

```{r}
#
#
#
#
#
#
#
#
#
#
#
#
#

```
<br />
<br />

# Read in Oscar Wilde's Flower of Love
We will try some of these regexp IRL. Let's use readLines to do this.
```{r}
#wilde <- readLines("flower.txt")
```


