---
title: "Week 2: Dataframes & Data Types"
author: "Jamie Reilly, Ph.D."
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: paged
    css: jamie_mod4.css
---

```{r setup, include=FALSE, cache=T}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', echo=TRUE, tidy=TRUE, message=F, warning=F, cache=T)
```

# Libraries 
```{r echo=T, message=F}
library(tidyverse)
library(janitor)
```

# Vectors
Uses the c() command. Contain all the same type of variable (integer, character, etc)
```{r}
emma <- c("a", "b", "c") #characters vs. strings
print(emma)
typeof(emma)
```

a new vector
```{r}
jamie <- seq(20, 25)
print(jamie)
typeof(jamie)
```

# Dataset 1
We'll read in and inspect a very simple dataframe (dat_simple) here.  
```{r}
easy <- read.csv("dat_simple.csv", header=T)  #read in data, header is first row
#remember no spaces, csv or text is best
```

Look at its structure and check for missing observations
```{r}
str(easy)
```

Print the whole dataframe
```{r}
easy
```

Examine just the first seven rows
```{r}
head(easy, n=7)
```

Examine just the last three rows
```{r}
tail(easy, n=6)
```

What is the smallest and largest shoe size?
```{r}
min(easy$Shoe) #amf
max(easy$Shoe) #amf
range(easy$Shoe) #ms
```

How many observations in my dataset?
```{r}
length(easy$IQ)
```

What is the average IQ?
```{r}
mean(easy$IQ)
summary(easy$IQ)
```

Scatterplot IQ by Shoe Size
```{r}
plot(easy$IQ, easy$Shoe, col="red")
```

# Using the %>% function
This very helpful function "chains" commands so that you do not have to
mutate(newvarname, operation)
newdata <- olddata %>% mutate(NewVariableName = SomethingMathy)
```{r}
print(easy)
easy2 <- easy %>% mutate(NewVar = IQ-Shoe)
median(easy$Shoe)
```

# ifelse function
ifelse(criteria, something, somethingelse)
```{r}
easy3 <- easy2 %>% mutate(BigFoot = ifelse(Shoe < 7, "tiny", "big"))
easy4 <- easy3 %>% mutate(LoHiGroup = ifelse(IQ <= 110, "lowfolk", "highfolk"))
print(easy4)
colnames(easy4)
```


# Subsetting columns and rows
dplyr filter and select commands. 
```{r}
little <- easy4 %>% select(1,2,6) #returns columns 1,2,4 to a smaller dataframe called 'little'
little2 <- little %>% filter(LoHiGroup == "highfolk") #just return the highfolk
print(little2)
```
<br />

# Scaling up in complexity
Now we will look at some much more complex data (dat_complex) using many of the same scripting commands you used with the simple data.  This dataset represents lexical characteristics for almost 80,000 English words. 

```{r}
hard <- read.csv("dat_complex.csv", header=T) %>% clean_names()
str(hard)
```

what is the average frequency per million words of an English word?
```{r}
hard %>% mean(brys_freqsubtlex, na.rm=T) #first way pipe -- Jamie get on this!
mean(hard$brys_freqsubtlex, na.rm=T) #second way $
```

what is the lowest imageability and highest imageability word in English?
```{r}
range(hard$gls_img)
```

what is the longest word in English?
```{r}

```

how many observations are in this dataset?
```{r}

```

how many missing observations are there for gls_img?
```{r}

```

what is the average letter length of a noun in English?
(hint use brys_pos as grouping variable)
```{r}

```

how many observations are there for each part of speech?
(hint use brys_pos and table function)
```{r}

```

create a grouping variable (abstract or concrete) using the median of concreteness as a cutpoint.
How many abstract words are there using this method?
(hint use mutate and ifelse to create a new variable with groups, then table that to get counts)
```{r}

```

create a scatterplot of age of acquisition by word frequency
(hint use brys_freqsubtlex as frequency and kup_aoa as age of acquisition)
```{r}

```

can you make any inferences about the relationship between the age at which a word is acquired and its frequency in English?
(hint just look at the scatterplot in the last question)
```{r}
#Write your answer here
```

create a smaller dataframe named "hard-small" that only reflects the first three columns of 'hard' and the first 1000 rows only.  Print the structure of hard-small

```{r}

```

append a new variable to 'hard' -- call it PhonMinLett - that reflects the number of phonemes minus the number of letters for each word
(hint use mutate with kup_nphon - kup_nlett)
```{r}

```

What word corresponds to row 8761?
```{r}

```


# Extra credit challenge!
Generate a histogram of word frequency in English
```{r}

```

Create a new variable meancentFreq that represents the mean centered values for word frequency?
Print its minimum and maximum and its structure

```{r}

```

Create a new variable zFreq that represents the z-transformed values for word frequency

```{r}

```

Append a sequence of numbers to "hard" from 1 to the end of the dataset that could be used as a counter variable -- call this variable ID_Num 
(hint - use seq_along with mutate)
```{r}

```


